Solution-aware vs global ReLU selection: \\ partial MILP strikes back for DNN verification

### 2025-07-23 ATVA 2025/Final_version/ATVA25.tex
ATVA25
authors

\documentclass[runningheads]{llncs}
\usepackage{hyperref}
\usepackage{url}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\pagestyle{plain}
\usepackage{threeparttable}
\input{math_commands.tex}
%\usepackage[latin9]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{cases}
\captionsetup{compatibility=false}
% \usepackage{esint}
\usepackage{array}
\usepackage{epstopdf}
\usepackage{placeins}
\usepackage{pgfplots}
\usepackage{url}
\usepackage{tikz}
\usepackage{calc}
\usepackage{array}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usetikzlibrary{positioning, arrows.meta,calc}

@@

\author{Yuke Liao\inst{1}\orcidID{0009-0004-3763-686X} \and
	Blaise Genest\inst{2,3}\orcidID{0000-0002-5758-1876} \and
	Kuldeep Meel\inst{4}\orcidID{0000-0001-9423-5270}
	\and
	Shaan Aryaman\inst{5}\orcidID{0000-0001-7576-0766}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{CNRS@CREATE, Singapore \email{yuke.liao@cnrsatcreate.sg} \and The corresponding author\\
	CNRS@CREATE, Singapore \email{blaise.genest@cnrsatcreate.sg}\and
	CNRS, France
	\email{blaise.genest@cnrs.fr}\\
 \and
	University of Toronto,  Toronto, Canada \\\email{meel@cs.toronto.edu}
	\and
	NYU Courant Institute of Mathematical Sciences, New York, USA\\ \email{aryaman.shaan@gmail.com}}
%
\maketitle              % typeset the header of the contribution
%

	\linenumbers
	\begin{abstract}
		Branch and Bound (BaB) is considered as the most efficient technique for DNN verification: it can propagate bounds over numerous branches, 
		to accurately approximate values a given neuron can take even in large DNNs, enabling formal verification of properties such as local robustness. Nevertheless, the number of branches grows {\em exponentially} with important variables, and there are complex instances for which the number of branches is too large to handle even using BaB. In these cases, providing more time to BaB is not efficient, as the number of branches treated is {\em linear} with the time-out. Such cases arise with verification-agnostic DNNs, non-local properties (e.g. global robustness, computing Lipschitz bound), etc. 
		%The fact that pure BaB is not that efficient for e.g. verification-agnostic (even very small) DNNs has been witnessed before. The workaround, e.g. in {\em refined} $\alpha,\beta$-CROWN, was to precompute very accurate bounds for the first few neurons of the DNN using a complete full MILP encoding. Non-surprisingly, this very slow technique does not scale but to small DNNs.
		%Indeed, one of its implementation, $\alpha,\beta$-CROWN has won the last 4 VNNcomp(etitions), as the DNN verifier with the best trade-off between accuracy and runtime. 
		%VNNcomp however is focusing on relatively easy verification problems.

@@

This opens up interesting future research directions, to verify global \cite{lipshitz}, \cite{sensing} rather than local (robustness) properties, which need very accurate methodology
and give rise to hard instances as the range of each neuron is no more local to a narrow neighborhood (and thus almost all ReLUs are unstable, with both modes possible).

@@

\newpage

\bigskip

\appendix

\input{proofsb}

\clearpage
\newpage

>>>>

\documentclass[runningheads]{llncs}
\usepackage{hyperref}
\usepackage{url}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\pagestyle{plain}
\usepackage{threeparttable}
\input{math_commands.tex}
%\usepackage[latin9]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{wrapfig}
%\usepackage{lineno}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{cases}
\captionsetup{compatibility=false}
% \usepackage{esint}
\usepackage{array}
\usepackage{epstopdf}
\usepackage{placeins}
\usepackage{pgfplots}
\usepackage{url}
\usepackage{tikz}
\usepackage{calc}
\usepackage{array}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage[misc]{ifsym}
\usetikzlibrary{positioning, arrows.meta,calc}

@@

\author{Yuke Liao \Letter \inst{1}\orcidID{0009-0004-3763-686X} \and
	Blaise Genest\inst{2,3}\orcidID{0000-0002-5758-1876} \and
	Kuldeep Meel\inst{4}\orcidID{0000-0001-9423-5270}
	\and
	Shaan Aryaman\inst{5}\orcidID{0000-0001-7576-0766}}
%
\authorrunning{Y. Liao, B. Genest, K. Meel, S. Aryaman}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{CNRS@CREATE, Singapore \email{yuke.liao@cnrsatcreate.sg} 
\and CNRS@CREATE \& IPAL, Singapore \email{blaise.genest@cnrsatcreate.sg}\and
	CNRS, IPAL, France \email{blaise.genest@cnrs.fr}\\
 \and
	University of Toronto,  Toronto, Canada \\\email{meel@cs.toronto.edu}
	\and
	NYU Courant Institute of Mathematical Sciences, New York, USA\\ \email{aryaman.shaan@gmail.com}}
%
\maketitle              % typeset the header of the contribution
%

	%\linenumbers

	\begin{abstract}
		Branch and Bound (BaB) is considered as the most efficient technique for DNN verification: it can propagate bounds over numerous branches, 
		to accurately approximate values a given neuron can take even in large DNNs, enabling formal verification of properties such as local robustness. Nevertheless, the number of branches grows {\em exponentially} with important variables, and there are complex instances for which the number of branches is too large to handle even using BaB. In these cases, providing more time to BaB is not efficient, as the number of branches treated is {\em linear} with the time-out. Such cases arise with verification-agnostic DNNs, non-local properties (e.g. global robustness, computing Lipschitz bound), etc. 
		%The fact that pure BaB is not that efficient for e.g. verification-agnostic (even very small) DNNs has been witnessed before. The workaround, e.g. in {\em refined} $\alpha,\beta$-CROWN, was to precompute very accurate bounds for the first few neurons of the DNN using a complete full MILP encoding. Non-surprisingly, this very slow technique does not scale but to small DNNs.
		%Indeed, one of its implementation, $\alpha,\beta$-CROWN has won the last 4 VNNcomp(etitions), as the DNN verifier with the best trade-off between accuracy and runtime. 
		%VNNcomp however is focusing on relatively easy verification problems.

@@

This opens up interesting future research directions, 
to verify global \cite{lipshitz}, rather than local (robustness) properties, which need very accurate methodology
and give rise to hard instances as the range of each neuron is no more local to a narrow neighborhood 
(most ReLUs are unstable, with both modes possible).

@@

\smallskip
{\bf Acknowledgement:} 
This research was conducted as part of the DesCartes program and 
was supported by the National Research Foundation, Prime Minister's Office, Singapore, 
under the Campus for Research Excellence and Technological Enterprise
(CREATE) program, and partially supported by ANR-23-PEIA-0006 SAIF.

### 2025-07-23 ATVA 2025/Final_version/Paper_noProof.tex
update draft of final version
update the draft of final version
authors

\documentclass[runningheads]{llncs}
\usepackage{hyperref}
\usepackage{url}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\pagestyle{plain}
\usepackage{threeparttable}
\input{math_commands.tex}
%\usepackage[latin9]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{cases}
\captionsetup{compatibility=false}
% \usepackage{esint}
\usepackage{array}
\usepackage{epstopdf}
\usepackage{placeins}
\usepackage{pgfplots}
\usepackage{url}
\usepackage{tikz}
\usepackage{calc}
\usepackage{array}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usetikzlibrary{positioning, arrows.meta,calc}

@@

\author{Yuke Liao\inst{1}\orcidID{0009-0004-3763-686X} \and
	Blaise Genest\inst{2,3}\orcidID{0000-0002-5758-1876} \and
	Kuldeep S. Meel\inst{4}\orcidID{0000-0001-9423-5270}
	\and
	Shaan Aryaman\inst{5}\orcidID{0000-0001-7576-0766}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{CNRS@CREATE, Singapore \email{yuke.liao@cnrsatcreate.sg} \and The corresponding author\\
	CNRS@CREATE, Singapore \email{blaise.genest@cnrsatcreate.sg}\and
	CNRS, France
	\email{blaise.genest@cnrs.fr}\\
 \and
	University of Toronto,  Toronto, Canada \\\email{meel@cs.toronto.edu}
	\and
	NYU Courant Institute of Mathematical Sciences, New York, USA\\ \email{aryaman.shaan@gmail.com}}
%
\maketitle              % typeset the header of the contribution
%

	\linenumbers
	\begin{abstract}
		Branch and Bound (BaB) is considered as the most efficient technique for DNN verification: it can propagate bounds over numerous branches, 
		to accurately approximate values a given neuron can take even in large DNNs, enabling formal verification of properties such as local robustness. Nevertheless, the number of branches grows {\em exponentially} with important variables, and there are complex instances for which the number of branches is too large to handle even using BaB. In these cases, providing more time to BaB is not efficient, as the number of branches treated is {\em linear} with the time-out. Such cases arise with verification-agnostic DNNs, non-local properties (e.g. global robustness, computing Lipschitz bound), etc. 
		%The fact that pure BaB is not that efficient for e.g. verification-agnostic (even very small) DNNs has been witnessed before. The workaround, e.g. in {\em refined} $\alpha,\beta$-CROWN, was to precompute very accurate bounds for the first few neurons of the DNN using a complete full MILP encoding. Non-surprisingly, this very slow technique does not scale but to small DNNs.
		%Indeed, one of its implementation, $\alpha,\beta$-CROWN has won the last 4 VNNcomp(etitions), as the DNN verifier with the best trade-off between accuracy and runtime. 
		%VNNcomp however is focusing on relatively easy verification problems.

@@

@@

\input{proofsb}

>>>>

\documentclass[runningheads]{llncs}
\usepackage{hyperref}
\usepackage{url}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\pagestyle{plain}
\usepackage{threeparttable}
\input{math_commands.tex}
%\usepackage[latin9]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{wrapfig}
%\usepackage{lineno}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{cases}
\captionsetup{compatibility=false}
% \usepackage{esint}
\usepackage{array}
\usepackage{epstopdf}
\usepackage{placeins}
\usepackage{pgfplots}
\usepackage{url}
\usepackage{tikz}
\usepackage{calc}
\usepackage{array}
\usepackage[misc]{ifsym}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usetikzlibrary{positioning, arrows.meta,calc}

@@

\author{Yuke Liao \Letter \inst{1}\orcidID{0009-0004-3763-686X} \and
	Blaise Genest\inst{2,3}\orcidID{0000-0002-5758-1876} \and
	Kuldeep Meel\inst{4}\orcidID{0000-0001-9423-5270}
	\and
	Shaan Aryaman\inst{5}\orcidID{0000-0001-7576-0766}}
%
\authorrunning{Y. Liao, B. Genest, K. Meel, S. Aryaman}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{CNRS@CREATE, Singapore \email{yuke.liao@cnrsatcreate.sg} 
\and CNRS@CREATE \& IPAL, Singapore \email{blaise.genest@cnrsatcreate.sg}\and
	CNRS, IPAL, France \email{blaise.genest@cnrs.fr}\\
 \and
	University of Toronto,  Toronto, Canada \\\email{meel@cs.toronto.edu}
	\and
	NYU Courant Institute of Mathematical Sciences, New York, USA\\ \email{aryaman.shaan@gmail.com}}
%
\maketitle              % typeset the header of the contribution
%

	%\linenumbers

	\begin{abstract}
		Branch and Bound (BaB) is considered as the most efficient technique for DNN verification: it can propagate bounds over numerous branches, 
		to accurately approximate values a given neuron can take even in large DNNs, enabling formal verification of properties such as local robustness. Nevertheless, the number of branches grows {\em exponentially} with important variables, and there are complex instances for which the number of branches is too large to handle even using BaB. In these cases, providing more time to BaB is not efficient, as the number of branches treated is {\em linear} with the time-out. Such cases arise with verification-agnostic DNNs, non-local properties (e.g. global robustness, computing Lipschitz bound), etc. 
		%The fact that pure BaB is not that efficient for e.g. verification-agnostic (even very small) DNNs has been witnessed before. The workaround, e.g. in {\em refined} $\alpha,\beta$-CROWN, was to precompute very accurate bounds for the first few neurons of the DNN using a complete full MILP encoding. Non-surprisingly, this very slow technique does not scale but to small DNNs.
		%Indeed, one of its implementation, $\alpha,\beta$-CROWN has won the last 4 VNNcomp(etitions), as the DNN verifier with the best trade-off between accuracy and runtime. 
		%VNNcomp however is focusing on relatively easy verification problems.

@@

\smallskip
{\bf Acknowledgement:} 
This research was conducted as part of the DesCartes program and 
was supported by the National Research Foundation, Prime Minister's Office, Singapore, 
under the Campus for Research Excellence and Technological Enterprise
(CREATE) program, and partially supported by ANR-23-PEIA-0006 SAIF.

@@

%
\input{proofsb}
